Tempus: Probabilistic Network Latency Verifier

### 2022-09-23 paper/main.tex
Added abstracts

\begin{document}
\begin{abstract}
    Test
\end{abstract}

@@

\section{Introduction}
As society develops further in the information age and gets more dependent on interconnected 
digital services, the task of making those services reliable is of paramount priority.
The network that underlies those services is no exception to this.

Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
routing information so that a packet can reach their intended destination efficiently, even in the 
event of a failure.
Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
It is hard for a network engineer to reason about whether their configured network fulfilled some 
intended property in various possible states of the network. 
Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
to help them in this task, namely, network verifier.

In recent years, there have been many work done in this area. 
Data plane verifiers work on a single snapshot of the forwarding table of a given network to 
verify a property. 
However, if we want to verify how a given routing protocol will react to a specific failure, 
control plane verifier might be used instead. 
A control plane verifier will efficiently search for failures that might cause change in the 
property in question.

Earlier control plane verifiers are formulated to answer questions about a given property 
deterministically, in other words, yes or no answer. 
For example, k-link failures.
While useful to some extent, this kind of models are sometimes too restrictive since network 
operators are usually able to tolerate failures for a small fraction of time, particularly when 
reasoning about compliance with Service Level Agreement (SLA) where, for example, a given network 
only needs to be operational 99.999\% of the time. 
This kind of probabilistic properties have been the focus of a more recent works
\cite{steffen2020probabilistic}.

The property itself can also be divided into two kinds. 
Quantitative properties of the network, like reachability, waypointing, and loop existence, are 
the common properties that are studied by most of the existing verifiers. 
More recent works, like QARC, has also explored qualitative properties like link bandwidth 
violation for a given traffic.

@@

\section{Proposed Approach}
We will start with the assertion that latency is a property that only make sense 
after connectivity between two nodes has been established. 
In other words, if two nodes in a network aren't connected (due to physical failure 
or ACL policy), then the latency between them will \textit{always} be infinite.
Because of this, we divide the problem of latency verification into two parts: 
verifying that two nodes are reachable (\textit{functional} property) and only if 
the functional property is fulfilled, we would verify whether the latency between 
two nodes fulfill some property (\textit{temporal} property)

@@

\section{Proposed Evaluation}
\begin{itemize}
    \item cross-validation with NetDice and DIRECT implementation in R
    \item performance evaluation with topology-zoo + dummy latency distribution
    \item performance evaluation with latency measurement data
\end{itemize}

>>>>

\begin{document}
\begin{abstract}
    To combat problems and bugs that a given network might have, network verifiers have emerged as 
    one of the promising solutions. 
    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
    various scenarios of network failures, such as reachability under k-link failure. 
    However, as modern networks evolved and performance need becomes more stringent -- often 
    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
    verifiers to also reason about quantitative performance properties. 
    Works in quantitative network verifiers that has arose in recent years mainly focuses on one 
    side of the network performance metric: bandwidth and load violation properties. 
    Questions about the other side of network performance metric, latency, were left unanswered.

    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
    a given temporal properties being true given latency distributions of individual links and 
    nodes in the network. 
    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
    that the average delay of packets traversing from a source to destination node is below $T$
    time unit -- by only adding a fraction of the overhead introduced by the qualitative verifier 
    its built upon.
\end{abstract}

@@

\section{Introduction}
Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
routing information so that a packet can reach their intended destination efficiently, even in the 
event of a failure.
Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
It is hard for a network engineer to reason about whether their configured network fulfilled some 
intended property in various possible states of the network. 
Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
to help them in this task, namely, network control plane verifier.

Earlier control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
questions about a given property deterministically. 
In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
expected to give a yes or no answer about whether a given property (e.g. two nodes are reachable) 
is satisfied in all of those cases. 
While useful to some extent, this kind of models are sometimes too restrictive since network 
operators are usually able to tolerate failures for a small fraction of time, particularly when 
reasoning about compliance with Service Level Agreement (SLA) where, for example, a given network 
only needs to be operational 99.999\% of the time. 
This kind of probabilistic properties have been the focus of a more recent works like NetDice 
\cite{steffen2020probabilistic}.

Aside from the state exploration, the property itself can also be divided into two kinds. 
Quantitative properties of the network, like reachability, waypointing, and loop existence, are 
the common properties that are studied by most of the existing verifiers. 
More recent works, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
properties like link bandwidth violation for a given traffic.

@@

\section{Proposed Approach}
We will start with the assertion that latency is a property that only make sense 
after connectivity between two nodes has been established. 
In other words, if two nodes in a network aren't connected (due to physical failure 
or ACL policy), then the latency between them will \textit{always} be infinite.
Because of this, we divide the problem of latency verification into two parts: 
verifying that two nodes are reachable (\textit{functional} property) and only if 
the functional property is fulfilled, we would verify whether the latency between 
two nodes fulfilled some property (\textit{temporal} property)

@@

% TODO: intro, reformalize graph formulation, citations
% Examples of workshop paper?
% Does a figure about topology graph and latency graph will be useful?
% Multimedia context on older networked system Klara Nahrstedts
% Graph

\section{Proposed Evaluation}
We want to evaluate two things: correctness and performance.

On the correctness side, we want to cross-validate the result of the components that make up 
the verifier, mainly the DIRECT algorithm, with the result of their publication. 
We also wanted to validate whether the result of our verifier is equivalent to another verifier 
that has lower fidelity (i.e. NetDice, that only verifies functional property) if we set the 
temporal property to be virtually unconstrained (e.g. bounded reachability with a very high 
$T$).

As for performance, we will evaluate the verifier by measuring the amount of additional 
overhead that the temporal verifier would introduce, measured by the amount of states that 
needs to be temporally explored and / or the amount of convolution operation that needs to be 
performed.
We will evaluate them using two main datasets: using topologies from the topology zoo \cite{knight2011internet} with a
realistic assumed temporal distribution and topologies that has node measurements that could 
be used to form a output queue distribution.
